\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{array}
\usepackage[numbers]{natbib}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[LO,LE]{Rominger, {\it et al.}}
\fancyhead[RO,RE]{Postdoc Mentoring Plan}

\usepackage[T1]{fontenc}
\usepackage{titling}
\setlength{\droptitle}{-3.5em}

\usepackage{wrapfig}

\usepackage{multirow}


%% make lists more compact
\usepackage{enumitem}
\setitemize{noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}
\setenumerate{noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}


%% make sections, subsections and paragraphs more compact
\makeatletter
\renewcommand\section{\@startsection{section}{1}{\z@}%
                                  {-1.5ex \@plus -1ex \@minus 0.2ex}%
                                  {0.1ex \@plus 0.2ex}%
                                  {\normalfont\Large\bfseries}}
\makeatother

\makeatletter
\renewcommand\subsection{\@startsection{subsection}{1}{\z@}%
                                  {-1.5ex \@plus -1ex \@minus 0.2ex}%
                                  {0.1ex \@plus 0.2ex}%
                                  {\normalfont\large\bfseries}}
\makeatletter

\makeatletter
\renewcommand\subsubsection{\@startsection{subsection}{1}{\z@}%
                                  {-1.5ex \@plus -1ex \@minus 0.2ex}%
                                  {0.1ex \@plus 0.2ex}%
                                  {\normalfont\bfseries}}
\makeatletter

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}
  {1.5ex \@plus 1ex \@minus .2ex}{-1em}
  {\normalfont\normalsize\it}
}
\makeatother


\title{Postdoc Mentoring Plan \vspace{-1.5ex}}
\author{}
\date{}


\begin{document}
\maketitle
\thispagestyle{fancy} 
\vspace{-6em}


Our two funded postdoc positions support PI Rominger and co-PI Owens.
While these researchers have already demonstrated maturity as scientists
and academics, we will nonetheless mentor Rominger and Owens following
the guidance of the National Academies of Science and Engineering on how
to enhance the postdoctoral experience \citep{NASpostdoc}.
Owens will be formally mentored by co-PIs Guralnick and Soltis, while
Rominger will be formally mentored by Guralnick and Hickerson, in
addition to his established career development mentor, Prof.~Jennifer
Dunne at the Santa Fe Institute, who serves in a capacity outside of
this grant proposal. All senior co-PIs will engage in informal mentoring
of the postdocs.

\paragraph{Co-development of an individual plan that defines the expectations
and goals of the postdoc:} At the beginning of the postdoc appointments,
the group will meet with Rominger and Owens to identify lead formal
mentors with who to develop plans that will provide the basis for
regular self-assessment, formal annual evaluation, and activities during
the project. This plan will be updated yearly via meetings between the
postdocs and their formal mentors.

\paragraph{Explicit acknowledgement that career development is an
  integral part of a post-doctoral position:} We will attend seminars,
workshops, skills training, and professional development events at the
Santa Fe Institute (SFI), the University of Florida (UF), and other
institutions. We specifically include opportunities covering
professional activities.  These occur as part of the non-hierarchical
culture at SFI, where Rominger participates in faculty meetings and
hiring committees, and within the UF Museum, which introduces Owens to
the service responsibilities of faculty, teaching effectiveness and
strategies for promoting student-centered learning. Both postdocs and
all senior co-PIs will engage in mentoring on how to find funding and
next-step career opportunities.

\paragraph{Strong support for postdocs' attendance at core meetings:}
Rominger and Owens will attend core disciplinary meetings such as the
Society for the Study of Evolution (SSE) and Ecological Society of
America (ESA), to promote and receive feedback on their work and gain
exposure to cutting-edge research. Rominger and Owens will also be
supported to take part in activities at iDigBio, which will provide an
unparalleled opportunity for them to get first-hand experience in 21st
century museum expertise.

\paragraph{Quarterly postdoctoral mentoring reviews to examine the status and
progress towards career goals:} These reviews will address concrete
project management and career development topics, e.g.~developing
rational timelines, major roadblocks, and identifying opportunities to
exercise leadership, etc. We will together provide Rominger and Owens
with an annual progress summary for potential use in reports to NSF.

\setlength{\bibsep}{5pt}
\bibliographystyle{grant}
\bibliography{postdocMentoring}

\end{document}

